---
title: "Homework 5"
subtitle: "Winter 2025"
author: "WRITE YOUR NAME HERE"
date: "`r Sys.Date()`"
output: pdf_document
header-includes:
    - \usepackage{amsmath, amssymb}
    - \usepackage{setspace} 
---

\onehalfspacing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)

```


* * * 

### Instructions

- This homework is due in Gradescope on Wednesday February 26 by midnight PST. There is a 15 minute grace period and submissions made during this time will not be marked as late. Any work submitted past this period is considered late.  

- Please answer the following questions in the order in which they are posed. 

- Don't forget to (i) make a local copy this document for your work and to (ii) knit the document frequently to make sure there are no compilation errors. 

- When you are done, download the PDF file as instructed in section and submit it in Gradescope. 

* * *


### Exercises 


1. (Blood type) In the US population, 85% have an agglutinating factor in their blood classifying them as Rh positive, while 15% lack the factor and are Rh negative. A medical researcher wants to analyze blood from a newborn Rh negative infant, so she examines the blood types of successive newborn infants until she finds an Rh negative infant. 


a. Let the random variable $X$ denote the number of Rh positive infants she needs to type before she finds the first negative case. Write the distribution of $X$. Be sure to state the values of any parameters. List the assumptions you need to make to answer this question.

b. How many Rh positive infants should she expect to type before she finds her first Rh negative? State the formula you are using, then calculate and print it in a code chunk  and then report your answer in a complete sentence in context.

c. What is the probability that she will type more Rh positive infants than expected? Give the expression for this probability here, do not calculate it yet. 

d. Calculate the probability from part c using your expression and also a built in R function for working with the distribution from part a. Be sure to echo your code chunk and report your answers.  
   
2. (Memoryless) Let $X$ be a geometric random variable with probability $\pi$. That is,
$$f(x) = \pi \cdot (1 - \pi)^{x}, \ \ x = 0, 1, 2 \dots$$ 

a. Show that for all non-negative integers  $x$ and $k$,
$$P(X \geq k + x | X \geq k) = P(X \geq x).$$ Be sure to show your steps.

b. Because of the result in part a, we say that the geometric distribution is a \emph{memoryless} distribution. Explain how this is an appropriate name for this property.

3. (Insecticide) A large number of insects are expected to be attracted to a variety of rose plant. A commercial insecticide is advertised as being 99% effective. Suppose 2,000 insects infest a rose garden where the insecticide has been applied and let $X$ denote the number of surviving insects.
  
a. What probability distribution might provide the most accurate model for this experiment? Be sure to give the reasons for why.


b. Calculate the probability that fewer than 10 insects survive, using the model from  part a. Be sure to show your code and output.

    ```{r 3b}
    
    
    ```

c. What other probability distribution might be more computationally convenient and would provide a good approximation? Be sure to state the name and values of any parameters.

d. Calculate  an approximation to the probability from part b using the model from part c; Be sure to show your code and output.

    ```{r 3d}
    
    
    ```
  


4. (Apple trees) From each of 6 trees in an apple orchard, 25 leaves are selected. On each of the leaves, the number of adult mites are counted. 

    \begin{table}[h]
    \centering
    \begin{tabular}{c|ccccccccc} \hline
    $x$ & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8+ \\
    count & 70 & 38 & 17 & 10 & 9 & 3 & 2 & 1 & 0 \\ \hline
    \end{tabular}
    \end{table}

a. Using R as a calculator, write code to find $\bar{x}$, the mean number of adult mites found. Be sure to show your code and the output!

    ```{r 4a}


    ```

b. Let $X$ denote the number of mites on a randomly selected leaf. In this question, you will consider two possible models for $X$:

 - $X \sim Pois(lambda = \bar{x})$ 
    
 - $X \sim Geom( prob = \frac{1}{1 + \bar{x}} )$
 
    where $\bar{x}$ is the number you calculated from part a. for the mean. 
  
    Write code in the code chunk below to make side-by-side plots of the probability histograms for each distribution. Be sure to show the code and the output, don't forget labels!

    *Hint*: For help with the code for side-by-side plots, look from slide 11 in the slidedeck Chapter 8.1. You will obviously need to modify this code for this example. 

    ```{r 4b}
    #add patchwork package in setup code chunk, NOT HERE!


    ```
    
    
c. Fill in the blanks in the code chunk below and then remove the `eval = FALSE` option. 


    ```{r 4c, eval = FALSE}
    
    obs_vs_pred <- tibble(
         x = 0:7,
         obs_prop =  ___ ,
         pred_pois = ___ ,
         pred_geom = ___
         
    )
    
    obs_vs_pred
    
    
    ```
    

d. Which model - the Poisson or Geometric -- seems to fit this data better? 
