---
title: "Homework 6"
subtitle: "Winter 2025"
author: "WRITE YOUR NAME HERE"
date: "`r Sys.Date()`"
output: pdf_document
header-includes:
    - \usepackage{amsmath, amssymb}
    - \usepackage{setspace} 
---

\onehalfspacing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)

```


* * * 

### Instructions

- This homework is due in Gradescope on Wednesday March 5 by midnight PST. There is a 15 minute grace period and submissions made during this time will not be marked as late. Any work submitted past this period is considered late.  

- Please answer the following questions in the order in which they are posed. 

- Don't forget to (i) make a local copy this document for your work and to (ii) knit the document frequently to make sure there are no compilation errors. 

- When you are done, download the PDF file as instructed in section and submit it in Gradescope. 

* * *


### Exercises 


1. (Flooding river) A river floods every year. Suppose that the low-water mark is set at 1 and the high-water mark $X$ has CDF
\begin{align*}
	F_X(x) &= P(X \leq x) \\
	       &= \left\{ \begin{array}{cc} 
		               0 & x < 1 \\
		               1 - 1/x^2 & 1 \leq x < \infty. \end{array} \right.
\end{align*}

a.  Find a PDF, $f(x)$ for $X$.

b. If the low-water mark is set to 0 and we use a unit of measurement that is $\frac{1}{10}$ of that given previously, the high-water mark becomes $Y=10\:(X-1).$ Find $P(Y \leq 1)$. 
   
    

2. (Phone calls) Suppose the duration of a phone call (in minutes), $X$, can be described probabilistically by the equation

	 $$P(X > x) = \frac{1}{2} e^{-x} + \frac{1}{2} e^{- \frac{x}{5}}, \ 0 \leq x < \infty$$
	   
   	
a. Write the CDF of $X$ by filling in the question marks in align environment below. (No explanation needed)

   \begin{align*}
   F(x) &= P(X \leq x) \\
   &= \left\{ \begin{array}{cc} 
                   ??? & x < 0\\
                   ??? & x \geq 0.
              \end{array} \right.
    \end{align*}
    
b. Find a PDF, $f(x)$  for $x$. Show your work mathematically. 

c. Find the probability that the call will last between 3 and 5 minutes. In other words, find $P(3 \leq X < 5)$. Show what you will be calculating first and then do the calculation in a code chunk and print the answer.

    *Hint:* You can just plug into the CDF and avoid integrations.
   
d. Find the median length of a phone call (in mins). Show what you are calculating mathematically first. Calculations should then be done in a code chunk and the answer reported.

    Hint: Please see  problem 3 from the week 6 Quiz section recording (Problem set 6). You want to find $q$ so that $P(X \leq q) = F(q) = 0.5$. You will need to use `uniroot` to solve this equation since it is non-linear in $q$. See 

3. Suppose $Y \sim Unif(-3,3)$. 

a. Write the PDF of $Y$. (Don't forget the range)

b. Find the probability that the roots of the quadratic equation in $x$
$$x^2 + x \cdot Y + (Y + 1)$$ are real. 

    Hint: The roots of the equation $ax^2 + b x + c=0$ are real if the discriminant $b^2-4ac \geq 0$.

4. The following ordered set of 27 Probability-values (from Kaati et al., Eur. J. Hum Genetics 2007) were the result of testing a hypothesis using many independent subgroups of a sample. If the hypothesis being tested is true, then the P-values all come from a uniform distribution on $[0,1)$.

    ```{r pvalues}
    pvalue_df <- tibble( 
      pval = c(0.01, 0.01, 0.02, 0.04, 0.04, 0.05, 0.07, 0.07, 
               0.10, 0.19, 0.24, 0.27, 0.34, 0.37, 0.44, 0.50, 
               0.53, 0.54, 0.55, 0.61, 0.70, 0.77, 0.80, 0.80, 
               0.82, 0.94, 0.99)
      )
    ```

      Using `ggplot`, make a histogram of the P-values. The superimpose the PDF of a $Unif(0,1)$ on it. What do you conclude about the hypothesis being tested?

     Hint: Please see problem 4 from the week 7 Quiz Section recording (Problem set 7) for sample code. Also when using the `breaks` argument, set the witdth of the bins to 1/6 for uniformity of grading.


