---
title: "Problem Section 5"
subtitle:  "Mon February 10"
graphics: yes
output: 
        pdf_document
header-includes:
    - \usepackage{amsmath, amssymb}
    - \usepackage{framed}\definecolor{shadecolor}{rgb}{0.949,0.949,0.949}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

* * * 
\begin{shaded}

\textbf{Learning Outcomes}

The problems are designed to build conceptual understanding and problem-solving skills. The emphasis is on learning to find, evaluate and build confidence.
The specific tasks include: 

   - Calculate expected value and variance using the PMF
     
   - Use Chebychev's inequality to make probability statements
  
   - Back up and support work with relevant explanations

 

\end{shaded}


* * *

### Exercises 

1. (Law of the Unconscious Probabilist) Weird Wally offers you the following choice. You may have a flat amount of $\frac{1}{3.5}$ dollars or you may roll a fair die and he will give you $\frac{1}{X}$ dollars where $X$ is the value on the roll. 
The PMF of $X$ is shown in the table below.

    \begin{table}[h]
    \centering
    \begin{tabular}{c|c|c|c|c|c|c} \hline
    $x$ & 1& 2 & 3 &4 & 5& 6 \\ \hline
    $f(x)$& $\frac{1}{6}$ & $\frac{1}{6}$ & $\frac{1}{6}$ &$\frac{1}{6}$ &$\frac{1}{6}$     &$\frac{1}{6}$ \\ \hline
    \end{tabular}
    \end{table}

    Which is the better deal? Calculate $E\left[ \frac{1}{X} \right]$ to decide.  (Hint: use the Law of the Unconscious Probabilist - Lemma 7.2)

2. (Mean and variance) 

a. Suppose a random variable $X$ has PMF as shown below. Find $E\left[X \right]$. Also calculate $SD\left[X \right]$.

    \begin{table}[h]
    \centering
    \begin{tabular}{|c|c|c|} \hline
      $x$ & -\$5 & \$170 \\ \hline 
      $f(x)$ & 37/38  & 1/38  \\ \hline
    \end{tabular}
    \end{table}

b. Suppose $X \sim Binom(n = 10, \pi = \frac{2}{3})$. What is the expected value of $Y = 3X - 4$?

c. If $X$ denotes a temperature of a randomly selected day recorded in degrees Fahrenheit, then $Y = \frac{5}{9} X - \frac{160}{9}$ is the corresponding temperature in degrees Celsius. If the standard deviation for $X$ is 15.7$^{\circ} F$, what is the standard deviation of $Y$?

    
3. (Chebychev) Suppose $X$ is a random variable with mean and variance both equal to 20. What can be said about $P( 0 <  X < 40)$?

    *Hint*: Chebychev's inequality says that 
    $$P(|X - 20| \geq k \sqrt{20} ) \leq \frac{1}{k^2}.$$ 
    What would you choose for $k$ here so you can say something about $P(0 < X < 40)$?
   

4. (Suppose we wish to generate $X \sim Binom(n=10, \pi=\frac{2}{3})$ subject to the constraint $X \leq 3$. Say we use the following naive algorithm to accomplish this task:

    - Generate an $x$ from a $Binom(10, \frac{2}{3})$
  
    - Accept the value $x$ if $x <= 3$. Otherwise reject it.
  
a. Calculate the acceptance probability. That is, what is the probability we will accept a value $x$ that is generated?

b. Put your calculation from part a to the test by generating 1,000 values from a $Binom(10, \frac{2}{3})$  and calculate how often you would generate an acceptable value. Bump up the number of simulations to 5,000. What do you notice?

    ```{r 4a}
    set.seed (131)  #for reproducibility
    ```


c. Define a new random variable $Y$ as the number of times we have to generate a binomial variable before we find an acceptable one. For example, if on our first try, we get an acceptable $x$, then $y=0$. Write the PMF of $Y$.  

<!---
c. How many $x$'s should you expect to reject? That is, find $E\left[Y \right]$.

    Hint: For any real number $r < 1$, we have the result 
    $$1 + r + r^2 + r^3 + \dots = \frac{1}{1-r}$$ 
--->