---
title: "Problem Section 6"
subtitle:  "Mon Feb 17, 2025"
graphics: yes
output: 
        pdf_document
header-includes:
    - \usepackage{amsmath, amssymb}
    - \usepackage{framed}\definecolor{shadecolor}{rgb}{0.949,0.949,0.949}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)  #for count function in problem 4
library(fastR2)     #for fumbles data

```

* * * 
\begin{shaded}

\textbf{Learning Outcomes}

The problems are designed to build conceptual understanding and problem-solving skills. The emphasis is on learning to find, evaluate and build confidence.
The specific tasks include: 

   - Use the geometric series to calculate probabilities
   
   - Use the Poisson model to calculate probabilities
   
   - Use the `uniroot` function to solve a non-linear equation
   
   - Fit a Poisson model to data
  
   - Back up and support work with relevant explanations

 
 

\end{shaded}


* * *

### Exercises 

1. (Geometric series) Andrea has found that she wins about 28% of the time when playing a solitaire game  on her computer. 

a. Calculate the probability that she plays 10 games before her first win.

b. Calculate the probability that she has to play at least 10 games before her first win.

c. What is the expected number of games she has to play before her first win? Find the probability that she loses more games than expected.


2.  (Poisson process) Suppose telephone calls arrive at a help line at the rate of two per minute. A Poisson process is assumed to provide a good model for the arrivals. In other words, if we let $X(t)$ denote the number of arrivals in an interval of length $t$ minutes, then $X(t) \sim Poisson(\lambda = 2t)$ and the PMF of $X(t)$ is give by
   \begin{align*}
   f(x) &= P(X(t) = x) = \frac{ e^{-2t} \:(2t)^x}{x!}  \ \ x = 0,1,2, \dots
   \end{align*}

    For each question below, write the probability you wish to calculate, then use R to do the calculation.

a. Calculate the probability that exactly five calls will arrive in the next 2 minutes.

b. Calculate the probability that the next two calls will arrive more than 5 minutes apart.



3. (Poisson) Suppose $X$, the number of chocolate chips in a certain type of cookie has a Poisson distribution with rate $\lambda$. We want the probability that a randomly chosen cookie has at least 2 chocolate chips to be greater than or equal to 0.99. Find the smallest value of $\lambda$ that ensures this probability.

   *Note* The equation will be non-linear in $\lambda$ and can only be solved numerically. For your part, just set up the equation which needs to be solved. You will learn how to use `uniroot` to find the $\lambda$ in section. 
   
   
4. Is the Poisson distribution a good fit for modeling the number of fumbles in NCAA football? The package **fastR2** contains a dataset called `Fumbles` which gives the number of fumbles by each NCAA FBS team for the first three weeks in November 2010. 

    The frequency distribution of the number of fumbles made in week 1 is shown below.
    
    ```{r message = F, warning = F}
    Fumbles %>% count(week1)
    
    ```
    
a. Using R as a calculator, write code in the code chunk below to find $\bar{x}$ the mean number of fumbles made in week 1.

    ```{r 4a}

    ```


b. Suppose we model these data using a Poisson distribution with rate $\lambda$ set equal to $\bar{x}$. What does this Poisson model predict for the probabilities of seeing 0 fumbles? How does this match with the fraction of times we actually observed "0" fumbles in week 1? Repeat for the remaining values.

    Again, use R as a calculator and write your code to do the calculations in the code chunk below. You may use the function `dois` for finding the probabilities which are predicted by the Poisson model.
    
    ```{r 4b}
 
    
    ```
    


