---
title: "Problem Section 7"
subtitle:  "Mon Feb 24, 2025"
graphics: yes
output: 
        pdf_document
header-includes:
    - \usepackage{amsmath, amssymb}
    - \usepackage{framed}\definecolor{shadecolor}{rgb}{0.949,0.949,0.949}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

* * * 
\begin{shaded}

\textbf{Learning Outcomes}

The problems are designed to build conceptual understanding and problem-solving skills. The emphasis is on learning to find, evaluate and build confidence.
The specific tasks include: 

   - Calculate density in a density histogram
   
   - Find probabilities using a PDF.
   
   - Derive the PDF from a CDF
   
   - Calculate probabilities in the uniform distribution
  
   - Back up and support work with relevant explanations

 

\end{shaded}


* * *

### Exercises 


1. Semester grades (expressed as a fraction of points) in a Physics for Poets class are decribed fairly well by the PDF
$$f(x) =  6x \cdot (1-x), \qquad 0 \leq x < 1.$$
Anyone earning less than 60% fails. Five housemates are among the students enrolled in the class. Suppose the random variable $X$ denotes the number (out of the 5 housemates) who fail the class.

a. What is the distribution of $X$? State the name as well as the values for the parameters. (What assumption are you making? Do you think it is valid here?)
b. The number of housemates who fail will be `___` give or take `___`. Fill in the first blank with $E\left[X \right]$ and the second blank with the $SD\left[X \right]$.

2. Suppose $X$ is a random variable with CDF
\begin{align*}
F(x) &= \left\{ \begin{array}{cc}
                      0 & x < 0 \\
                      1 - e^{-x^{3/2}} & x \geq 0
                  \end{array}
                  \right.
\end{align*}

    Determine $f(x)$, the PDF of $X$.


3. Suppose $X$ is uniformly distributed over $[0,1)$. That is
$$f(x) = 1 \ \ \ 0 \leq x < 1.$$ 
    Show that $$P(-\ln(X) \leq 5) = 1 - e^{-5}$$ 
   where $\ln(X)$ denotes denotes the natural logarithm of $X$.
   
   \begin{align*}
      P(X \geq e^{-5}) &= 1-P(X<e^{-5})\\
      &=1-\int_0^{e^{-5}}dx\\
      &=1- x \Big\rvert_0^{e^{-5}}\\
      &=1-e^{-5}
   \end{align*}
   
   
   
4. Suppose we have reason to believe that these
forty xi ’s may be a random sample from a uniform probability function defined over the interval $[20, 70)$ - that is,
$$f(x) = \frac{1}{50} \ \ \ 20 \leq x < 70.$$

    ```{r unif-data, echo = TRUE, message =TRUE, warning = TRUE, collapse = TRUE}

    sample_data <- tibble( 
          xval  = c( 33.8, 62.6, 42.3, 62.9, 32.9, 58.9, 60.8, 49.1, 
                     42.6, 59.8, 41.6, 54.5, 40.5, 30.3, 22.4, 25.0, 
                     59.2, 67.5, 64.1, 59.3, 24.9, 22.3, 69.7, 41.2, 
                     64.5, 33.4, 39.0, 53.1, 21.6, 46.0, 28.1, 68.7, 
                     27.6, 57.6, 54.8, 48.9, 68.4, 38.4, 69.0, 46.6 ) )
    ```

    In the following exercise, we will learn to make a histogram of these data and then superimpose the uniform PDF on the histogram to assess the fit. 
    
a. To begin, fill in the blanks in the code below to create a frequency histogram of the data. Once the blanks are filled, remove the `eval = F` chunk option to see the graph.

    ```{r freq-hist, echo = T, message = F, warning = F}

    ggplot() +
       geom_histogram(data = sample_data,
                 mapping =aes(x = xval ),
                 breaks = seq(20, 70, 10),
                 alpha = 0.5,
                 color = "purple",
                 fill = "gold") +
       labs(x = "xval",
            title = "Frequency Histogram") 

    ```

b. Before we can add the PDF of the uniform density to the graph, the y axis of the histogram must be re-scaled from a count (or frequency) to a density 
$$density = \frac{frequency}{n \times width}$$
so that the sum of the areas of the blocks equals 1. (The $n$ represents the total number of observations in the dataset.)

    What is the area of a block if the $y$ value is the density for that block? Why do the areas of the blocks add to 1 in a density histogram?
    
    The are of a block if we are plotting a density function is
    $$area = density \times width = \frac{frquency}{n}$$.
    
    The area of the blocks in a density histogram  add to 1 because we are adding 
    up the proportion of observations which fall in each bin ($\frac{frequency}{n}$.


c. Fill in the blanks in the following code to create a density histogram of `xval` and then superimpose the PDF of a $Unif(20,70)$ on it. Remove the `eval = F` chunk option after filling in the blanks.

    ```{r density-hist, echo = T, message = F, warning = F}

    ggplot() +
       geom_histogram(data = sample_data,
                 mapping =aes(x = xval, 
                              y = after_stat(density)),
                 breaks = seq(20, 70, 10),
                 alpha = 0.5,
                 color = "purple",
                 fill = "gold") +
      geom_function( fun = dunif, args = list( min = 20, max = 70) ) +
       labs(x = "xval",
            title = "Density Histogram with Uniform pdf") 

    ```



